Molecular Gas Dynamics of the lensed wet-merger RXSJ1131-1123 at $z$=0.654

### 2016-06-21 WRITEUP/ms.tex
shuffled down kinamtics discusion from obs. section, updated Mgas decomposition (didn't correct for lensing in some channels previously)

\begin{figure}[!htbp]
\includegraphics[trim=3 5 0 15, clip, width=0.242\textwidth]{../Figures/{F555WCO21_mom0_single.invertedgray}.eps}
\includegraphics[trim=3 5 0 0, clip, width=0.23\textwidth]{../Figures/F555W_REDBLUE.eps}
\caption{Left: an overlay of the velocity-integrated \bco emission on the {\it HST} $V$--band (F555W)
image. Right: contours are color-coded to represent the red and blue wings of the emission.
The contours start at 3$\sigma$ and increment at steps of
$\pm$3$\sigma$, where $\sigma$\,=\,0.3 mJy beam\pmOne. The crosses denote the
location of the foreground galaxy at $z$\,=\,0.295.
\label{fig:CO21mom}}
\end{figure}

The emission of the red component coincides with the Einstein Ring seen in the
optical image, with most of its apparent flux originating from the lensed arc
in the southeast, whereas the blue component is predominately coming from
solely the lensed arc shown in \Fig{CO21mom}. To illustrate this, we show the
channel maps of 21.5\,\kms width and spatial spectra of 1\farcs5 resolution in
\Fig{chanmap} and \Fig{spatialSpec}, respectively. The figures
show that emission
is present to the west, peaking toward the lensing arc (black crosses in
\Fig{chanmap}) in the red wing, and shifts to the east with decreasing velocity
(blue wing).
This indicates extended emission in the source plane, where
emission from different kinematic components is separated by different amounts
relative to the caustics, leading to the observed asymmetric
line profile. A detailed analysis on differential lensing of \bco is presented
in \Sec{differential}, where we present our lens model.

@@

\subsection{Origin of the \bco Emission} \label{sec:origin} %DONE
Evidently, the CO emission is co-spatial with the lensed optical emission in
the background AGN host galaxy.
The question is then whether the optically
faint companion also contributes to the CO flux and thus contains a
non-negligible amount of molecular gas.
Indeed, based on our lens model presented in \Sec{lensmodel}, we find it
highly likely that part of the CO emission originates from
the companion. We thus classify RXJ1131 as a ``wet-wet'' merger.
However, we cannot quantify this with a mass ratio given the
spatial resolution, nor can we rule out the possibility that
the ``companion'' is instead a gas component/tidal feature ripped out
from previous interaction.
In the subsequent sections, we will interpret
the this additional source component as gas emission
in a nearby galaxy (i.e. a ``wet-wet'' merger), as suggested by previous studies of RXJ1131 (C06, B08).

\subsection{\bco Kinematics} % DONE
A clear velocity gradient and a high
velocity dispersion ($\gtrsim$400\,\kms) near the central region
is seen in \Fig{CO21highO}. While beam smearing is inevitably the
dominant factor in the observed velocity dispersion
at the spatial resolution of this data, the exceedingly
high velocity dispersion may hint
at potential perturbations from the AGN, or internal turbulence due to
interactions with the companion, and/or instability due to the large gas
content. A velocity gradient across the source plane in \Fig{model} is also
found in our lens model. This further supports the idea of a
kinematically-ordered galaxy, but its emission has been lensed differentially.
In this scenario, RXJ1131 is a disrupted disk galaxy hosting an optically
bright quasar and is in the process of merging.
We will defer the discussion of its implications to \Sec{dynamics},
where we analyze the dynamics of the system in the source plane.

@@

Assuming the spatial extents between \bco and \cco are similar and therefore
magnified by similar amounts, the line intensities
correspond to a brightness temperature ratio of
$r_{\rm 32}$\,=\,T$_{\small \cco}$$/$T$_{\small \bco}$\,=\,0.66\,$\pm$\,0.41.

@@

\subsubsection{Interpretation of the source-plane morphology} \label{sec:caveat}
The reconstructed source locations in \Fig{model} demonstrate
an intrinsic velocity gradient across the source plane, which is
indicative of a disk-like galaxy.
Additional support to the disk conjecture
can be found in the double-horned line profile (\Fig{CO21spec})
and the observed (image plane) velocity field (\Fig{CO21highO}). Furthermore,
C06 also find that the reconstructed source plane emission in optical-NIR
is best-reproduced using a $n$\,=\,1 Sersic profile.
We thus interpret RXJ1131 to be a disk galaxy.

% the companion
One other interesting result from our lens model is that a better fit is
found for the red-most channel if we add a second source component (see
top left panel in \Fig{model}). This is consistent with previous results
reported by \citet{Brewer08a}, who find an optically faint companion
(component F in their paper) $\sim$2.4\,kpc away from the AGN host galaxy,
and C06, who find evidence for an interacting galaxy near RXJ1131.
Spatially, the red velocity component of the CO emission
also coincides with this component F. It is therefore evident that we
detect \bco emission in the companion galaxy.
To quantify the type of merger (major v.s. minor) conventionally with a mass
ratio, we decompose the total line flux into two components:
one from RXJ1131 and the other from its companion.
Since the companion is only detected in the red-most channel, we
use the best-fit flux densities and magnification factors obtained
from the model of this channel to derive the gas mass in the companion.
Without direct \rot{1}{0} measurements, we derive \rot{1}{0} intensity
assuming a brightness temperature ratio of $r_{\rm 21}$\,=\,1, which was chosen
solely to facilitate a comparison with other U/LIRGs. % also ULIRGs are found to be thermalized up to Jup = 3 or 4
For the same reason, we adopt a luminosity-to-mass conversion factor of
\alphaco\,=\,0.8 (K \kms pc$^2$)\pmOne. This yields a gas mass
$M_{\rm gas}$ of (1.68\pmm0.68)\,\E{10} \Msun. 
%statistical unc. only (i.e. draw from from distributions of on mag. factor & intrinsic flux)
For the molecular gas mass in RXJ1131, we derive
its intrinsic line flux over the FWZI linewidth
using the respective magnification
factors listed in \Tab{model}, which to
first order takes into account effect of differential lensing.
This yields $I_{\small \bco}$\,=\,5.41$\pm$2.71 Jy\,\kms,
where the uncertainty includes those on
the magnification factors.
Adopting the same brightness temperature ratio and \alphaco\ as
used for the companion, this corresponds to a gas mass of
$M_{\rm gas}$\,=\,(3.17\pmm1.59)\,\E{10} \Msun.
The implies a gas mass ratio of $\sim$2:1.
Therefore, the system is a gas-rich major merger according to
the classification scheme commonly used in literature.

@@

\subsubsection{Differential Lensing} \label{sec:differential} %DONE
\input{table_magFactor}
Similar to previous studies of RXJ1131, where
differential lensing across {\it HST}
$H$-, $V$-, and $I$-band has been detected with a
magnification factor decreasing from 10.9 to 7.8 (C06),
the highly asymmetric \bco line profile suggests that
differential lensing is also non-negligible for the CO emission, whereby
causing the redshifted emission to be apparently much brighter than the
blueshifted component.
Such an effect is consistent with the source plane
positions relative to the caustics in \Fig{model}; the red wing emission mainly originates near the cusp
in the caustic, whereas the blue wing emission is located beyond the caustics.
From our best-fit lens models, the magnification factor varies from 8.7 to 3.1 across the \bco line (\Tab{model}).

%If we estimate the intrinsic I:
%red I =
%blue wing I =

>>>>

\begin{figure}[!htbp]
\includegraphics[trim=0 10 15 0, clip, width=0.222\textwidth]{../Figures/{F555WCO21_mom0_single.invertedgray}.eps}
\includegraphics[trim=5 -10 0 0, clip, width=0.232\textwidth]{../Figures/F555W_REDBLUE.eps}
\caption{Left: an overlay of the velocity-integrated \bco emission on the {\it HST} $V$--band (F555W)
image. Right: contours are color-coded to represent the red and blue wings of the emission.
The contours start at 3$\sigma$ and increment at steps of
$\pm$3$\sigma$, where $\sigma$\,=\,0.3 mJy beam\pmOne. The crosses denote the
location of the foreground galaxy at $z$\,=\,0.295.
\label{fig:CO21mom}}
\end{figure}

@@

@@

Assuming the spatial extents between \bco and \cco are similar and therefore
magnified by the same amount, the line intensities
correspond to a brightness temperature ratio of
$r_{\rm 32}$\,=\,T$_{\small \cco}$$/$T$_{\small \bco}$\,=\,0.66\,$\pm$\,0.41.

@@

\subsubsection{Interpretation of the source plane morphology} \label{sec:caveat}
The reconstructed source locations in \Fig{model} demonstrate
an intrinsic velocity gradient across the source plane, which is
indicative of a disk-like galaxy.
Additional support to the disk conjecture
can be found in the double-horned line profile (\Fig{CO21spec})
and the observed (image plane) velocity field (\Fig{CO21highO}). Furthermore,
C06 also find that the reconstructed source plane emission in optical-NIR
is best-reproduced using a $n$\,=\,1 Sersic profile.
We thus interpret RXJ1131 to be a disk galaxy.

% the companion
One other interesting result from our lens model is that a better fit is
found for the red-most channel if we add a second source component (see
top left panel in \Fig{model}). This is consistent with previous results
reported by \citet{Brewer08a}, who find an optically faint companion
(component F in their paper) $\sim$2.4\,kpc away from the AGN host galaxy,
and C06, who find evidence for an interacting galaxy near RXJ1131.
Spatially, the red velocity component of the CO emission
also coincides with this component F. It is therefore evident that we
detect \bco emission in the companion galaxy.
To quantify the type of merger (major v.s. minor) conventionally with a mass
ratio, we decompose the total line flux into two components:
one from RXJ1131 and the other from its companion.
Since the companion is only detected in the red-most channel, we
use the best-fit flux densities and magnification factors obtained
from the model of this channel to derive the intrinsic gas mass in the companion.
Without direct \rot{1}{0} measurements, we derive \rot{1}{0} intensity
assuming a brightness temperature ratio of $r_{\rm 21}$\,=\,1, which was chosen
solely to facilitate a comparison with other U/LIRGs. 
% also ULIRGs are found to be thermalized up to Jup = 3 or 4, but not needed to say here
For the same reason, we adopt a luminosity-to-mass conversion factor of
\alphaco\,=\,0.8 (K \kms pc$^2$)\pmOne. This yields a gas mass
$M_{\rm gas}$ of (1.92\pmm0.09)\,\E{9} \Msun. 
%statistical unc. only (i.e. draw from distributions of on mag. factor & intrinsic flux)
For the molecular gas mass in RXJ1131, we derive
its intrinsic line flux over the FWZI linewidth
using the respective magnification
factors listed in \Tab{model}, which to
first order takes into account effect of differential lensing.
This yields $I_{\small \bco}$\,=\,2.93\pmm0.70 Jy\,\kms,
where the uncertainty includes those on
the magnification factors.
Adopting the same brightness temperature ratio and \alphaco\ as
used for the companion, this corresponds to a gas mass of
$M_{\rm gas}$\,=\,(1.38\pmm0.33)\,\E{10} \Msun,
implying a gas mass ratio of $\sim$7:1 between RXJ1131 and its companion.
We thus classify the system to be a ``wet-wet'' minor merger according to
the classification scheme/convention commonly used in literature.

@@

\subsubsection{Spatial extent and differential lensing} \label{sec:differential}
In the image plane shown in  \Fig{CO21mom}, the redshifted component is 
cospatial with the Einstein Ring seen in the
optical image, with most of its apparent flux originating from the lensed arc
in the southeast, whereas the blue component is predominately coming from
solely the lensed arc. To further illustrate this, we show the
channel maps of 21.5\,\kms width and a spatial spectra map of 1\farcs5 resolution in
\Fig{chanmap} and \Fig{spatialSpec}, respectively. The figures
show that emission
is present to the west, peaking toward the lensing arc (black crosses in
\Fig{chanmap}) in the red wing, and shifts to the east with decreasing velocity
(blue wing).
This is consistent with the source plane positions in our models and 
suggests that the 
overall CO emitting region is extended.

\input{table_magFactor}
Similar to previous studies of RXJ1131, where
differential lensing across {\it HST}
$H$-, $V$-, and $I$-band has been detected with a
magnification factor decreasing from 10.9 to 7.8 (C06),
the highly asymmetric \bco line profile suggests that
differential lensing is also non-negligible for CO,
causing the redshifted emission to be apparently much brighter than the
blueshifted component and the asymmetric line profile.   
To first order, this can be explained by the difference in magnification factor which
varies from 8.7 to 3.1 across the \bco line (\Tab{model}).  % also because of a contribution from the companion in the observed, lensed spectrum
Such an effect (difference in $\mu_{\rm L}$ across channels) is consistent with the source plane
positions relative to the caustics in \Fig{model}, where the red wing 
emission mainly originates near the cusp 
of the caustic and the blue wing emission is located beyond the caustics.
In fact, the intrinsic line flux of the redshifted and 
blueshifted emission in RXJ1131 (after subtracting a contribution from the companion)
is $I_{\small \bco}$\,=\,1.26\pmm 0.23 Jy\,\kms and 1.25\pmm0.23 Jy\,\kms, respectively, 
implying an intrinsically symmetric line profile.

\subsection{\bco Kinematics} % DONE
A clear velocity gradient and a high
velocity dispersion ($\gtrsim$400\,\kms) near the central region
is seen in \Fig{CO21highO}. While beam smearing is inevitably the
dominant factor in the observed velocity dispersion
at the spatial resolution of this data, the exceedingly
high velocity dispersion may hint
at potential perturbations from the AGN, or internal turbulence due to
interactions with the companion, and/or instability due to the large gas
content. A velocity gradient across the source plane in \Fig{model} is also
found in our lens model. This further supports the idea of a
kinematically-ordered galaxy, but its emission has been lensed differentially.
In this scenario, RXJ1131 is a disrupted disk galaxy hosting an optically
bright quasar and is in the process of merging.
